---
title: "MATH 216 Homework 3"
author: "Andrea Narciso"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(Quandl)
library(forcats)
library(knitr)
```



## Admistrative:

Please indicate

* Who you collaborated with:
* Roughly how much time you spent on this HW so far:
* The URL of the RPubs published URL [here](https://www.nhl.com/).
* What gave you the most trouble:
* Any comments you have:




## Question 1:

We will use a logistic regression model to predict sex. Our metric to rate how well our
model performs will be:

$$
\frac{1}{n}\sum_{i=1}^{n}I(y_i = \widehat{y}_i)
$$

where $I(A)$ is the *indicator function* that is equal to 1 if condition $A$
holds, 0 otherwise. So

* Say user $i$ **is** female, then $y_i=1$
* Say we **predict** user $i$ is female, then $\widehat{y}_i=1$
* In this case $I(y_i =\widehat{y}_i)=1$. 

So what the above formula is reporting is the proportion of users' sex we
correctly predicted.

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
# To keep this exercise simple, let's remove the only 3 users (0.005% of users)
# who did not list a height, define the outcome variable, and add an ID variable
# to distinguish the users
profiles <- read_csv(file="profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")
```



#### a)

Define:

* A *training* set `training` of 2997 users (5% of users). We will train the 
logistic regression model to predict gender using this data. Since we want to 
train the model to tell who is female and who is not, we use the outcome
variable `is_female`.
* A *test* set `test` of the remaining 56,946 users (95% of users). We will test
how good our trained model is using this data. So at first, we will pretend we
don't know the outcome variable `is_female`. We use the above model to make a
prediction of sex for all 56,946 test users, then we use the `is_female` outcome
to rate how well we performed.
* Be sure to incorporate all the insight your garnered in your EDA in HW-2.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

has_kids <- c("has a kid", "has kids")
has_kids_no_more <- c("has kids, but doesn&rsquo;t want more",
        "has a kid, but doesn&rsquo;t want more")
has_kids_wants_more <- c("has a kid, and wants more","has kids, and wants more")
has_kids_might_want_more <- c("has kids, and might want more", "has a kid, and might want more")
anti_kids <- c("doesn&rsquo;t have kids, and doesn&rsquo;t want any", "doesn&rsquo;t want kids", "doesn&rsquo;t have kids")
# "might want kids", "wants kids", "doesn&rsquo;t have kids, but wants them", "doesn&rsquo;t have kids, but might want them", NA

profiles <- profiles %>%
  select(-contains("essay")) %>% 
  mutate(kids = ifelse(is.na(offspring), "no_ans",
                       ifelse(offspring %in% anti_kids, "anti_kids",
                              ifelse(offspring %in% has_kids, "has_kids",
                                     ifelse(offspring %in% has_kids_no_more, "has_kids_no_more",
                                            ifelse(offspring %in% has_kids_might_want_more, "has_kids_might_want_more",
                                                   ifelse(offspring == "might want kids", "might_want",
                                                          ifelse(offspring == "wants kids", "wants",
                                                                 ifelse(offspring == "doesn&rsquo;t have kids, but wants them", "no_kids_but_wants", "no_kids_might_want")))))))),
         income_cat = ifelse(income == -1, "no_reply", 
                         ifelse(income < 50000, "low_income",
                         ifelse((income >=50000) & (income <= 100000), 
                                "mid_income", "high_income"))),
         job = ifelse(is.na(job), "no_report", job)) %>% 
  select(job, is_female, kids, income_cat)
profiles
# NOTE TO SELF: CAN ALSO fct_recode(col_title, "new" = "old", ...)

training <- profiles[sample(nrow(profiles), 2997), ]
  
test <- profiles[sample(nrow(profiles), 56946), ]

```



#### b)

Train the logistic regression model to predict sex. i.e. fit a logistic
regression model to the `training` data. Assign this model to an R object called
`predict_sex_model`, then rate how well the model performs on the `training` data.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

predict_sex_model <- glm(is_female ~ job + kids + income_cat, data=training, family="binomial")


```



#### c)

Take `predict_sex_model` and apply it to the `test` data and make a prediction 
for each users' sex, then rate how well the model performs on the `test` data.

**Hint**: What do you think `predict(predict_sex_model, newdata=test,
type="response")` does? The help file is located in `?predict.glm`

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

# temp_test <- test %>% 
#   select(job, kids, income_cat)
# temp_test

train_check <- training %>% 
  mutate(p_hat = predict(predict_sex_model, type = "response"), 
         probably_f = ifelse(p_hat > 0.5, 1, 0),
         correct = ifelse(probably_f == is_female, 1, 0)) %>% 
  summarise(avg_correct_train = round(mean(correct), 3))
train_check

test_check <- test %>% 
  mutate(p_hat = predict(predict_sex_model, newdata=test, type="response"),
         probably_f = ifelse(p_hat > 0.5, 1, 0),
         correct = ifelse(probably_f == is_female, 1, 0)) %>% 
  summarise(avg_correct_test = round(mean(correct), 3))
test_check

```



#### d)

Did the model perform better on the `training` data or the `test` data? Why
do you think that is?

The average number of correct predictions of the training data was 0.657, which is greater than that of the test data, which is 0.644. This could be because the model is created using the training data, and thus, is better at predicting the training data, than the new, test data.





## Question 2:

We want to compare the volatility of 

* [Bitcoin](https://www.quandl.com/data/BAVERAGE/USD) prices
* [Gold](https://www.quandl.com/data/BUNDESBANK/BBK01_WT5511) prices

Let our measure of volatility be the relative change from day-to-day in price. 
Let the reference currency be US dollars. Analyze these results and provide
insight to a foreign currency exchanger.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

bitcoin <- Quandl("BAVERAGE/USD") %>%
  tbl_df() %>% 
  rename(daily_avg = `24h Average`,
         total_vol = `Total Volume`) %>% 
  mutate(relative_price_change = daily_avg-lag(daily_avg, 1))

gold <- Quandl("BUNDESBANK/BBK01_WT5511") %>% 
  tbl_df() %>% 
  mutate(relative_price_change = Value - lag(Value, 1))

# Graphs all dates
gold_bitcoin_graph <- ggplot() +
  geom_line(data=gold, aes(x=Date, y = relative_price_change, col = 'gold')) +
  geom_line(data=bitcoin, aes(x = Date, y = relative_price_change, col = 'bitcoin')) +
  labs(y = "Value", title = "Relative Daily Price Change of Gold and Bitcoin Over Time")
gold_bitcoin_graph

# Graphs starting from 2010
gold_bitcoin_graph_2010 <- gold_bitcoin_graph +
  coord_cartesian(xlim=c(dmy("01-07-2010"),dmy("31-10-2016")),
                  ylim=c(-100, 150))
gold_bitcoin_graph_2010

```

Since bitcoin started in 2016, the value of the bitcoin has shown to change to a greater degree on a daily basis than the value of gold, indicating that bitcoin is a more volatile frequency than gold. Thus, as a foreign currency exchanger, I would prefer to use gold, as it a more stable and reliable currency.




## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


Using the Reed College jukebox data, what are the top 10 artists played during
the "graveyard shift" during the academic year? Define

* the "graveyard shift" as midnight to 8am
* the academic year as September through May (inclusive)

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

jukebox_clean <- jukebox %>% 
  mutate(date_time = parse_date_time(date_time, "a b d HMS Y"), 
         month = month(date_time),
         hour = hour(date_time)) %>% 
  filter(!between(month, 6, 8), between(hour, 0, 8)) %>% 
  group_by(artist) %>% 
  tally() %>% 
  rename("num_played" = n) %>% 
  arrange(desc(num_played)) %>% 
  head(10) %>% 
  kable()
jukebox_clean

```





